Collaborative Infrastructure for Test-Driven\\Scientific Model Validation

### 2014-03-15 sciunit-icse14/sciunit-icse14.tex
ICSE14: Final draft rtpos
ICSE14: Fixed some final typos

\begin{abstract}
One of the pillars of the modern scientific method is \emph{model validation}: comparing a scientific model's predictions against empirical observations. Today, a scientist demonstrates the validity of a model by making an argument in a paper and submitting it for peer review, a process comparable to \emph{code review} in software engineering. While human review helps to ensure that contributions meet high-level goals, software engineers typically supplement it with \emph{unit testing} to get a more complete picture of the status of a project, particularly when it is complex and involves many contributors.

We argue that a similar test-driven methodology would be valuable to scientific communities as they seek to  validate increasingly complex models against growing repositories of empirical data. Scientific communities differ from software communities in several key ways, however. In this paper, we introduce \emph{SciUnit}, a framework for test-driven scientific model validation and outline how SciUnit, supported by new and existing collaborative infrastructure, could be  integrated into the modern scientific workflow.
\end{abstract}
\category{D.2.5}{Software Engineering}{Testing and Debugging}%[complexity measures, performance measures]
\terms{Documentation, Measurement, Standardization}
\keywords{unit testing, model validation, cyberinfrastructure}

@@

Quantitative models are today validated by \emph{peer review}. For a model to be accepted by a scientific community, its advocates must submit a paper that describes how it works and provides evidence that it makes more accurate predictions than previous models (or that it makes a desirable tradeoff  between accuracy and complexity) \cite{box1987empirical}. Other members of the relevant community are then tasked with ensuring that validity was measured properly and that all relevant data and competing models were adequately considered, drawing on knowledge of statistical methods and the prior literature. Publishing is a primary motivator for most scientists \cite{howison2011scientific}.

@@

Code review can be quite effective \cite{codereview}, but this requires that developers  expend considerable effort \cite{kemerer2009impact}. 
%Code review is also most effective for resolving  issues related to software architecture. 
Most large development teams thus supplement code reviews with more {automated} approaches to verification and validation, the most widely-used  of which is \emph{unit testing} \cite{beck2003}. In brief, unit tests are functions that check that the behavior of a single component satisfies a single functional  criterion. 
%A suite of such tests can be considered a partial specification of the program or component under development.  
A suite of such tests complements code review by making it easier to answer questions like these:
\begin{enumerate}
\item What functionality is a component expected to have?
\item What functionality has been adequately implemented? What remains to be done?
%\item How much progress does an individual code contribution make? 
\item Does a candidate code contribution cause \emph{regressions} in other parts of a program?
\end{enumerate}
Scientists ask analagous questions:
\begin{enumerate}
\item What is a model's scope and how is validity measured?
\item Which observations are already explained by existing models? What are the best models of a particular quantity?  What data has yet to be explained?
%\item How much better is a new model than the state of the art? What are the open problems in my field?
\item What affect do new observations have on the validity of previously published models? Can new models explain previously published data?
\end{enumerate}
But while software engineers can rely on a program's test suite, scientists today must extract this information from a body of scientific papers. This is increasingly difficult. Each paper is narrowly focused, often considering just one model or dataset, and is frozen in time, so it does not consider the latest models, data or statistical methods. Discovering, precisely characterizing and comparing models to discover the state of the art and find open modeling problems can require an encyclopedic knowledge of the literature, as can finding all data relevant to a model. Senior scientists often attempt to fill this need by publishing review papers, but in many areas, the number of publications generated every year can be overwhelming \cite{jinha_article_2010}, and comprehensive reviews of a particular area are published infrequently. Statisticians often complain that scientists are not following best practices and that community standards evolve too slowly because a canonical paper or popular review used outdated methods. Furthermore, if the literature simply doesn't address an important question of validity, a researcher might need to reimplement an existing model in order to evaluate it.

@@

To begin to address these challenges, we will introduce a lightweight and flexible validation testing and framework called \textit{SciUnit} in Sec. 2. Many of these challenges have to do with coordination between scientists. To begin to address this, we then describe a community workflow based on widely-adopted social coding tools (here, Github) and a lightweight community portal called \textit{SciDash}, in Sec. \ref{scidash}. The overall goal of this work is to help scientists generate and examine tables like the one central to Figure 1, where the relative validity of a set of models having a common scope can be determined by examining scores produced by a suite of validation tests constructed from experimental data. 
We discuss adoption strategies and directions for future research into scientific model validation practices in Sec. \ref{discussion}.
%Anecdotally, one of the greatest fears of new scientists is that a problem that they are working on has already been solved and published.

@@

This illustrates a key difference between unit testing, which would simply produce a boolean result, and our conception of scientific validation testing (challenge 1). A score class must induce an ordering, so that a table like that shown in Figure 1 can be sorted along its columns, and it can  specify a normalization scheme so the cells can be color-coded.  %It also makes it easier to cache predictions generated by large-scale models so that they can be used by multiple tests.

Te \verbx{judge} method of a test can be invoked to compute a score for a single model: 
\vspace{-4px}
\begin{python}
score = saturn_position_test.judge(lin_saturn_model)
\end{python}
\vspace{-5px}
This method proceeds by first checking that the provided model implements all required capabilities before calling the \verbx{generate_prediction} method followed immediately by the \verbx{compute_score} method. A reference to the test,   model, observation, prediction and other related data the test provides (none here) are available as attributes of the score.

@@

Social coding tools like Github allow each scientist to work on their own fork of a repository like \verbx{saturnunit} where they can examine the consequences of their own contributions on  the tests and models that have previously been developed. For example, an experimentalist who has gathered new data can instantiate a new test or refine an existing one  with more accurate observations. As long as the interface between the test and the model remains the same, the scores for prior models can be recomputed entirely automatically, requiring no coordination with model developers (challenge 4). When a modeler develops a new model, she can   validate it against all the tests it is capable of taking. If a statistician develops a better goodness-of-fit metric, it too can existing capabilities, and so a comprehensive score table can be generated easily, as discussed above. If a scientist is not interested in certain tests, she can locally remove them to recalibrate her view of the relative merits of different models (challenge 5).

Once satisfied with their contributions to the repository, these scientists might submit a \emph{pull request} to the central repository (either separately or together with the submission of a conventional paper). Anonymous reviewers tasked by the ``editors'' of the repository would then subject the pull request to appropriate peer review before committing it to the ``academic record''. A model that performs well across tests in a collaboratively developed suite could more directly and believably claim (e.g. to reviewers) to be a coherent, valid model of, for example, Saturn's motion.

@@

To help scientists discover and coordinate control over these repositories, we are developing a collaborative portal called \textit{SciDash} (\url{http://scidash.org/}). It is essentially a wiki pointing scientists to the common and suite repositories designated as ``primary'' by early adopters in their research areas (who serve as \emph{de facto} initial editors as well). SciDash also supports loading IPython notebooks containing score tables stored in these repositories into the browser directly, avoiding the need to clone the repository locally. To modify the suite (e.g. to add the model a scientist is developing to it), a scientist can simply fork the repository by clicking a button at the top of this view. We are exploring ways of loading forks directly into a cloud-based IPython notebook system, so that the complexities of Git and installation of Python and the packages required by the repository are left behind the scenes (challenge 4). SciDash discovers public forks of repositories that it is indexing and lists them to help scientists get a better sense of who is working on the problems they are interested in and what partial results are publicly available, fostering collaboration.

As with many pieces of cyberinfrastructure, early adopters may benefit less than late adopters due to network effects. To incentivize adoption, we look to an increasingly common scientific practice: modeling competitions. Many research communities use these to periodically determine the state of the art, resulting in a form of review paper. Funding sources often include small monetary incentives.
%For example, the quantitative single neuron modeling competition operated over the course of two years  \cite{qsnmc}. 
   SciDash could be used to run these kinds of competitions on a continuous basis atop less \textit{ad hoc} infrastructure (and thus require less effort and central planning from organizers) and we are actively organizing such competitions.

@@

\vspace{5px}
\section{Discussion \& Future Directions}\label{discussion}
Instead of attempting to use an existing general-purpose testing infrastructure, we have designed what is essentially a domain-specific language that captures the scientific model validation process in more familiar terms (challenge 6). SciUnit can be used by individual scientists to organize their workflows and also within a loosely collaborative workflow, mediated by a social coding tool like Github and an easy-to-use portal, SciDash, based on the IPython system \cite{perez2007ipython}.

@@

While we discuss a simple example based on planetary movement here, we have applied this framework to more realistic problems in neurobiology in collaboration with a large-scale informatics project, OpenSourceBrain, whose developers are enthusiastic about precisely characterizing the scope and validity of the dozens of models they currently maintain (case study in preparation). The tools we describe could also be used for a variety of other scientific disciplines, including within software engineering to validate quantitative models of, for example, software cost or reusability.

This paper identifies potential  synergies between software development and model validation practices, and provides a proof of concept to argue that a test-driven development methodology could help improve scientific workflows. The modern scientific model validation process has not yet been extensively studied by  the research community. Precisely characterizing how scientists today attempt to answer the questions in the introduction, and evaluating the effectiveness of tools like the ones we propose, are likely fruitful avenues for future empirical research.

@@

\section{ACKNOWLEDGEMENTS}
We thank Sharon Crook, Shreejoy Tripathy and Padraig Gleeson for their many helpful discussions.  
This work was supported in part by NIMH grant R01MH081905. 
The content is solely the responsibility of the authors and does not necessarily represent the official views of the NIH.
\vspace{5px}
% conference papers do not normally have an appendix

>>>>

\begin{abstract}
One of the pillars of the modern scientific method is \emph{model validation}: comparing a scientific model's predictions against empirical observations. Today, a scientist demonstrates the validity of a model by making an argument in a paper and submitting it for peer review, a process comparable to \emph{code review} in software engineering. While human review helps to ensure that contributions meet high-level goals, software engineers typically supplement it with \emph{unit testing} to get a more complete picture of the status of a project.%, particularly one involving many contributors.

We argue that a similar test-driven methodology would be valuable to scientific communities as they seek to  validate increasingly complex models against growing repositories of empirical data. Scientific communities differ from software communities in several key ways, however. In this paper, we introduce \emph{SciUnit}, a framework for test-driven scientific model validation, and outline how, supported by new and existing collaborative infrastructure, it could be  integrated into the modern scientific process.
\end{abstract}
\category{D.2.5}{Software Engineering}{Testing and Debugging}%[complexity measures, performance measures]
\terms{Documentation, Measurement, Standardization}
\keywords{unit testing, model validation, cyberinfrastructure}

@@

Today, quantitative models are validated by \emph{peer review}. For a model to be accepted by a scientific community, its advocates must submit a paper that describes how it works and provides evidence that it makes more accurate predictions than previous models (or that it makes a desirable tradeoff  between accuracy and complexity) \cite{box1987empirical}. Other members of the relevant community are then tasked with ensuring that validity was measured correctly and that all relevant data and competing models were adequately considered, drawing on knowledge of statistical methods and the prior literature. Publishing is a primary motivator for most scientists \cite{howison2011scientific}.

@@

Code review can be quite effective \cite{codereview}, but this requires that developers  expend considerable effort \cite{kemerer2009impact}. 
%Code review is also most effective for resolving  issues related to software architecture. 
Most large development teams thus supplement code reviews with more {automated} approaches to verification and validation, the most widely-used  of which is \emph{unit testing} \cite{beck2003}. In brief, unit tests are functions that check that the behavior of a single component satisfies a single functional  criterion. 
%A suite of such tests can be considered a partial specification of the program or component under development.  
A suite of such tests complements code review by making it easier to answer questions like these:
\begin{enumerate}
\item What functionality is a component expected to have?
\item What functionality has been adequately implemented? What remains to be done?
%\item How much progress does an individual code contribution make? 
\item Does a candidate code contribution cause \emph{regressions} in other parts of a program?
\end{enumerate}
Scientists ask analagous questions:
\begin{enumerate}
\item What is a model's scope and how is validity measured?
\item Which observations are already explained by existing models? What are the best models of a particular quantity?  What data has yet to be explained?
%\item How much better is a new model than the state of the art? What are the open problems in my field?
\item What affect do new observations have on the validity of previously published models? Can new models explain previously published data?
\end{enumerate}
But while software engineers can rely on a program's test suite, scientists today must extract this information from a body of scientific papers. This is increasingly difficult. Each paper is narrowly focused, often considering just one model or dataset, and is frozen in time, so it does not consider the latest models, data or statistical methods. Discovering, precisely characterizing and comparing models to determine the state of the art and find open modeling problems can require an encyclopedic knowledge of the literature, as can finding all data relevant to a model. Senior scientists often attempt to fill these needs by publishing review papers, but in many areas, the number of publications generated every year can be overwhelming \cite{jinha_article_2010}, and comprehensive reviews of a particular area are published infrequently. Statisticians often complain that scientists are not following best practices and that community standards evolve too slowly because a canonical paper or popular review relied on outdated methods. Furthermore, if the literature simply doesn't address an important question of validity (e.g. if space was limited), a researcher might need to reimplement an existing model from scratch to attempt to address it.

@@

To begin to address these challenges, we will introduce a lightweight and flexible validation testing and framework called \textit{SciUnit} in Sec. 2. Many of these challenges have to do with coordination between scientists. To begin to address this, we then describe a community workflow based on widely-adopted social coding tools (here, Github) and a lightweight community portal called \textit{SciDash}, in Sec. \ref{scidash}. The overall goal of this work is to help scientists generate tables like the one central to Figure 1, where the relative validity of a set of models having a common scope can be determined by examining scores produced by a suite of validation tests constructed from experimental data. 
We discuss adoption strategies and conclude in Sec. \ref{discussion} with directions for future research into scientific model validation practices.
%Anecdotally, one of the greatest fears of new scientists is that a problem that they are working on has already been solved and published.

@@

This illustrates a key difference between unit testing, which would simply produce a boolean result, and our conception of scientific validation testing (challenge 1). A score class must induce an ordering, so that a table like that shown in Figure 1 can be sorted along its columns. A score class can also specify a normalization scheme so the cells of the table can be color-coded.  %It also makes it easier to cache predictions generated by large-scale models so that they can be used by multiple tests.

The \verbx{judge} method of a test can be invoked to compute a score for a single model: 
\vspace{-4px}
\begin{python}
score = saturn_position_test.judge(lin_saturn_model)
\end{python}
\vspace{-5px}
This method first checks that the model implements all required capabilities, then calls the \verbx{generate_prediction} method, then passes the prediction to the \verbx{compute_score} method. A reference to the test,   model, observation, prediction and other related data the test provides (none here) are added as attributes of the score before it is returned.

@@

Social coding tools like Github allow each scientist to work on their own fork of a repository like \verbx{saturnunit} where they can examine the consequences of their own contributions on  the tests and models that have previously been developed. For example, an experimentalist who has gathered new data can instantiate a new test or refine an existing one  with more accurate observations. As long as the interface between the test and the model remains the same, the scores for prior models can be recomputed entirely automatically, requiring no coordination with model developers (challenge 4). When a modeler develops a new model, she can   validate it against all the tests it is capable of taking. If a statistician develops a better goodness-of-fit metric, she too can use existing capabilities and indeed inherit prediction extraction logic directly, as discussed above. If a scientist is not interested in certain tests, she can locally remove them to recalibrate her view of the relative merits of different models (challenge 5).

If the changes to the repository appear to represent a contribution, these scientists might submit a \emph{pull request} to the central repository (either separately or together with the submission of a conventional paper). Anonymous reviewers tasked by the ``editors'' of the repository would then subject the pull request to appropriate peer review before committing it to the ``permanent record''. A model that performs well across tests in a collaboratively developed suite could more objectively claim (e.g. to reviewers) to be a coherent, valid model of, for example, Saturn's motion.

@@

To help scientists discover and coordinate development of these repositories, we are developing a collaborative portal called \textit{SciDash} (\url{http://scidash.org/}). It is essentially a wiki pointing scientists to the repositories designated as ``primary'' by early adopters in their research areas (who may serve as \emph{de facto} initial editors as well). SciDash also supports loading IPython notebooks containing score tables stored in these repositories into the browser, avoiding the need to clone the repository locally. To modify the suite (e.g. to add the model a scientist is developing to it), a scientist can simply fork the repository by clicking a button at the top of this view. We are exploring ways of loading forks directly into a cloud-based IPython notebook system, so that the complexities of Git, Python setup and package management can be left behind the scenes (challenge 6). SciDash lists public forks of repositories that it is indexing to help scientists get a better sense of who is working on the problems they are interested in and what partial results are publicly available, fostering collaboration.

As with many pieces of cyberinfrastructure, early adopters may benefit less than late adopters due to network effects. To incentivize adoption, we look to an increasingly common scientific practice: modeling competitions. Many research communities use these to periodically determine the state of the art, resulting in a form of review paper. Funding sources often include small monetary incentives.
%For example, the quantitative single neuron modeling competition operated over the course of two years  \cite{qsnmc}. 
   SciDash could be used to run these kinds of competitions on a continuous basis atop less \textit{ad hoc} infrastructure (and thus require less effort and central planning from organizers) and we are actively recruiting organizers for such competitions.

@@

\vspace{5px}
\section{Discussion \& Future Directions}\label{discussion}
Instead of attempting to use an existing general-purpose testing infrastructure, we have designed what is essentially a domain-specific language that captures the scientific model validation process in more familiar terms (challenge 6). SciUnit can be used by individual scientists to organize their workflows using IPython notebooks \cite{perez2007ipython} and also within a loosely collaborative workflow, mediated by a social coding tool like Github and a simple portal, SciDash.

@@

While we discuss a simple example based on planetary movement here, we have applied this framework to more realistic problems in neurobiology in collaboration with a large-scale informatics project, OpenSourceBrain, whose developers are enthusiastic about precisely characterizing the scope and validity of the dozens of models they currently maintain (case study in preparation). The tools we describe could also be used for a variety of other scientific disciplines, including within software engineering to validate predictive models of, for example, development time or code reuse.

This paper identifies potential  synergies between software development and model validation practices, and sketches a sociotechnical solution to several challenges to argue that a test-driven development methodology could help improve the scientific process. The scientific model validation process has not yet been extensively studied, so precisely characterizing how scientists attempt to answer the questions in the introduction today, and evaluating how effective tools like the ones we propose are in improving their ability to do so, are likely fruitful avenues for future empirical research.

@@

\section{ACKNOWLEDGEMENTS}
We thank Sharon Crook, Shreejoy Tripathy and Padraig Gleeson for their many helpful discussions, and the anonymous referees and several friends for their useful comments.  
RG was supported in part by NIMH grant R01MH081905. 
The content is solely the responsibility of the authors and does not necessarily represent the official views of the NIH.
\vspace{5px}
% conference papers do not normally have an appendix
